---
title: "Simple R Objects"
author: "Daffa"
date: "2023-03-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simple R Objects

### Types of Data

* **class** - dividing variables into logical, numerical, character or factor
* **type** - gives more information about how the variables are **stored**, and divides variables into logical, integer, double or character

R stores categorical variables as integers, but their true nature is revealed by their class, which is factor or ordered factor

```{r}
typeof(4)
typeof(3.141)
typeof("hello")
class(4)
class(3.14)
class("hello")
```

These functions work on anything in R - even on themselves

```{r}
typeof(sqrt)
class(sqrt)
class(class)
```

### Division with integers
```{r}
#simple division
25/7

#The number of complete times 7 fits into 25 - an integer value
25%/%7

#25 modulo 7 - the remainder after taking 25%%7 7s away from 25
25%%7
```

### Rounding numbers

```{r}
#round to the nearest whole number 
round(163.141593)

#round to the nearest while number
round(163.141593,0)

#round to one decimal place
round(163.141593,1)

#round to the nearest multipe of 10
round(163.141593,-1)

#round up
ceiling(163.141593)

#round down
floor(163.141593)
```

### R Objects

#### Numeric Objects

```{r}
x <- 2
y <- 6.4
x2 <- x^2
fred <- 2*x+5
wilma <- x+y
a.simple.fraction <- (x+1)/(y+2)

a.simple.fraction
```

### Logical objects

```{r}

2 < 3

6 == (2+3)

6 == (3+3)

is.logical(2)
bb <- TRUE
is.logical(bb)
is.numeric(4.1)
is.logical(a.simple.fraction)
is.logical(2>4)
is.logical(fred)
is.numeric(wilma)
is.numeric(fred)
is.character(wilma)

class(a.simple.fraction)
typeof(a.simple.fraction)

class(wilma)
typeof(wilma)
```
